Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2019-06-10 thesis/chapters/ch_chapter4.tex
typos and small edits

\begin{figure*}
\centering
\includegraphics[width=0.45\linewidth]{figures/ch4/AGB1_enrichment_evolution_average_CGM}
\includegraphics[width=0.45\linewidth]{figures/ch4/NSNS1_enrichment_evolution_average_CGM} \\
\includegraphics[width=0.45\linewidth]{figures/ch4/SNE1_enrichment_evolution_average_CGM}
\includegraphics[width=0.45\linewidth]{figures/ch4/HNE1_enrichment_evolution_average_CGM}
\caption{Time evolution of the fraction of metals in each phase of the ISM (colored lines; CNM: dark blue, WNM: light blue, WIM: orange, HIM: red), the galaxy's disk (black, dashed), and the CGM (black, solid) as averaged across all events at a given $E_{\rm ej}$. The fractions are all normalized to the total amount of metals initially injected in each event, and the individual ISM phases sum to the black, dashed line. }
\label{ch4:fig:ISM_CGM}
\end{figure*}

We summarize the source-averaged results for all enrichment events in runs \runone E46, \runone E49, \runone E50, \runone E51, and \runone E52 in Figure~\ref{ch4:fig:ISM_CGM} by showing the fraction of source metals contained within each phase of the ISM (colored lines, which sum to the black dashed line), and the CGM (black, solid). These plots show a clear, immediate trend across $E_{\rm ej}$ for all lines in the figure, but the differences are most striking for the ISM (black, dashed), CGM (black, solid) lines. In Chapter~\ref{ch:chapter3}
% \citep{Emerick2019}
we demonstrated that there was a significant difference in metal ejection fraction, $f_{\rm ej}$, for metals from AGB sources as compared to metals released in SNe. Here, we confirm that this is driven predominantly by differences in the energy of the events, and not necessarily when or where they occur in the ISM. The clear trend in Figure~\ref{ch4:fig:ISM_CGM} shows that events with higher $E_{\rm ej}$ are much more readily ejected from the disk of the galaxy, while events with lower $E_{\rm ej}$ couple poorly to the same galactic outflows.
%\textit{statement on how immediate enrichment is after events}
% By 10~Myr after the enrichment events,
The high-energy events rapidly converge on a peak $f_{\rm ej}$ within about 20~Myr of the event, with only a gradual increase towards the end of the 150~Myr as the metals in the ISM are swept up in additional outflows. The lower energy events evolve more gradually. The lowest energy event, corresponding to AGB winds, reaches $f_{\rm ej} \sim$~0.68, which increases to 0.87 for the 10$^{51}$~erg events, and 0.95 for the 10$^{52}$~erg event. To emphasize these differences, we plot each line separately in the left panel of Figure~\ref{ch4:fig:CGM_CNM}

>>>>

\begin{figure*}
\centering
\includegraphics[width=0.45\linewidth]{figures/ch4/AGB1_enrichment_evolution_average_CGM}
\includegraphics[width=0.45\linewidth]{figures/ch4/NSNS1_enrichment_evolution_average_CGM} \\
\includegraphics[width=0.45\linewidth]{figures/ch4/SNE1_enrichment_evolution_average_CGM}
\includegraphics[width=0.45\linewidth]{figures/ch4/HNE1_enrichment_evolution_average_CGM}
\caption{Time evolution of the fraction of metals in each phase of the ISM (colored lines; CNM: dark blue, WNM: light blue, WIM: orange, HIM: red), the galaxy's disk (black, dashed), and the CGM (black, solid) as averaged across all events at a given $E_{\rm ej}$. The fractions are all normalized to the total amount of metals initially injected in each event. The individual ISM phases sum to the black, dashed line. }
\label{ch4:fig:ISM_CGM}
\end{figure*}

We summarize the source-averaged results for all enrichment events in runs \runone E46, \runone E49, \runone E50, \runone E51, and \runone E52 in Figure~\ref{ch4:fig:ISM_CGM} by showing the fraction of source metals contained within each phase of the ISM (colored lines, which sum to the black dashed line), and the CGM (black, solid). These plots show a clear, immediate trend across $E_{\rm ej}$ for all lines in the figure, but the differences are most striking for the ISM (black, dashed) and CGM (black, solid) lines. In Chapter~\ref{ch:chapter3}
% \citep{Emerick2019}
we demonstrated that there was a significant difference in metal ejection fraction, $f_{\rm ej}$, for metals from AGB sources as compared to metals released in SNe. Here, we confirm that this is driven predominantly by differences in the energy of the events, and not necessarily when or where they occur in the ISM. The clear trend in Figure~\ref{ch4:fig:ISM_CGM} shows that events with higher $E_{\rm ej}$ are much more readily ejected from the disk of the galaxy, while events with lower $E_{\rm ej}$ couple poorly to the same galactic outflows.
%\textit{statement on how immediate enrichment is after events}
% By 10~Myr after the enrichment events,
The high-energy events rapidly converge on a peak $f_{\rm ej}$ within about 20~Myr of the event, with only a gradual increase towards the end of the 150~Myr as the metals in the ISM are swept up in additional outflows. The lower energy events evolve more gradually. The lowest energy event, corresponding to AGB winds, reaches $f_{\rm ej} \sim$~0.68, which increases to 0.87 for the 10$^{51}$~erg events, and 0.95 for the 10$^{52}$~erg event. To emphasize these differences, we plot each line separately in the left panel of Figure~\ref{ch4:fig:CGM_CNM}

### 2019-06-10 thesis/chapters/ch_conclusion.tex
typos and small edits

This \dissertation has focused on building an understanding of galactic chemical evolution with the point of view that the same feedback processes that regulate star formation and drive galactic winds in galaxies also play a fundamental role in their metal enrichment. It is understood that stellar feedback plays an important role in driving metals out into the CGM and beyond in galaxies, but the exact coupling of metals to galactic winds is not well understood. This is in part because how metals mix within a multi-phase ISM to eventually be swept up in galactic outflows or mix within cold, star forming gas is not well constrained. Even more, how metals from the full range of nucleosynthetic sources, including AGB winds, stellar winds, core collapse and Type Ia supernovae, and NS-NS mergers, participate in these processes -- mixing and galactic winds -- differently has only recently been examined in efforts to better understand results from the ever increasing availability of high-quality stellar abundance measurements of stars in our own Milky Way and in nearby dwarf galaxies. In order to build a complete picture of metal evolution in the Universe, we have developed a new model for star formation, stellar feedback, and chemical enrichment to follow these processes in simulations of isolated dwarf galaxies. Using this model, we have made critical advancements in understanding both how stellar feedback governs the evolution of low mass dwarf galaxies and how individual metals are distributed in (and beyond) these galaxies over time. We summarize these results below, and conclude with a discussion of future research.

@@

In ongoing work we are extending these methods by implementing them in cosmological simulations of UFD galaxy evolution. Our detailed model tracking the feedback and yields from individual stars will build substantially upon recent works investigating metal enrichment at high redshift and in UFDs and their progenitors \citep[e.g.][]{Ritter2015,Jeon2017,Corlies2018,Wheeler2018,Agertz2019}. This work includes additional improvements to the simulations presented in this \Dissertation, including a model for Pop III star formation and chemical enrichment adapted from \cite{Wise2012a}, a new IMF sampling scheme that allows us to follow very low mass stars ($M_* \lesssim 2$~M$_{\odot}$) -- which have no significant feedback or metal enrichment, but are important long-lived tracers of galactic chemical enrichment -- aggregated together into a single particle during each star formation event, improvements to the numerical implementation to ease memory consumption in these significantly larger simulations, and new metal tracer fields to follow the contribution of AGB, core collapse SNe, Type Ia SNe, and PopIII stars to the total metal content of each cell independently.

Finally, as part of this \dissertation we additionally produced a significant set of companion simulations at $2 \times$ lower resolution (3.6~pc instead of 1.8~pc) turning on and off each independent feedback process (ionizing radiation, radiation pressure, LW and FUV radiation, and supernovae). With these runs, we sought to build upon recent works examining the various effects of each stellar feedback process in driving the evolution of low mass dwarf galaxies \citep{Hu2016,Hu2017,Hu2018,Forbes2016}. In future work we will examine these completed runs in more detail to better understand the complex, interconnected role multi-channel stellar feedback plays in galactic evolution. With our simulations, we can place greater emphasis on how each feedback channel drives the mixing and ejection of individual metals within and beyond low mass dwarf galaxies.

@@

In addition, this work can be readily extended to further investigate how feedback governs the evolution of dwarf galaxies. We have already produced a suite of comparison simulations at a factor two lower resolution (3.6~pc) than our fiducial resolution, turning on and off each individual feedback process to understand the significance and characterize the mechanism by which each operates to regulate star formation, drive galactic winds, and govern chemical evolution. The feedback tested here includes supernovae, stellar ionizing radiation, photoelectric heating, Lyman-Werner radiation, and radiation pressure. The suite contains simulations including ``all-but-one" of each process, in addition to several combinations where multiple effects are excluded, and a set where the effectiveness of radiation pressure is artificially increased and decreased. Comparing the evolution of these dwarfs to our fiducial simulations presented in this \dissertation will improve our understanding of these processes, building upon previous works that have conducted similar studies in more massive dwarf galaxies \citep{Hu2016,Hu2017,Forbes2016,Hopkins2018}.

>>>>

This \dissertation has focused on building an understanding of galactic chemical evolution with the point of view that the same feedback processes that regulate star formation and drive galactic winds in galaxies also play a fundamental role in their metal enrichment. It is understood that stellar feedback plays an important role in driving metals out into the CGM around galaxies and beyond, but the exact coupling of metals to galactic winds is not well understood. This is in part because how metals mix within a multi-phase ISM to eventually be swept up in galactic outflows or mix within cold, star forming gas is not well constrained. Even more, how metals from the full range of nucleosynthetic sources, including AGB winds, stellar winds, core collapse and Type Ia supernovae, and NS-NS mergers, participate in these processes -- mixing and galactic winds -- differently has only recently been examined in efforts to better understand results from the ever increasing availability of high-quality stellar abundance measurements of stars in our own Milky Way and in nearby dwarf galaxies. In order to build a complete picture of metal evolution in the Universe, we have developed a new model for star formation, stellar feedback, and chemical enrichment to follow these processes in simulations of isolated dwarf galaxies. Using this model, we have made critical advancements in understanding both how stellar feedback governs the evolution of low mass dwarf galaxies and how individual metals are distributed in (and beyond) these galaxies over time. We summarize these results below, and conclude with a discussion of future research.

@@

%In ongoing work we are extending these methods by implementing them in cosmological simulations of UFD galaxy evolution. Our detailed model tracking the feedback and yields from individual stars will build substantially upon recent works investigating metal enrichment at high redshift and in UFDs and their progenitors \citep[e.g.][]{Ritter2015,Jeon2017,Corlies2018,Wheeler2018,Agertz2019}. This work includes additional improvements to the simulations presented in this \Dissertation, including a model for Pop III star formation and chemical enrichment adapted from \cite{Wise2012a}, a new IMF sampling scheme that allows us to follow very low mass stars ($M_* \lesssim 2$~M$_{\odot}$) -- which have no significant feedback or metal enrichment, but are important long-lived tracers of galactic chemical enrichment -- aggregated together into a single particle during each star formation event, improvements to the numerical implementation to ease memory consumption in these significantly larger simulations, and new metal tracer fields to follow the contribution of AGB, core collapse SNe, Type Ia SNe, and PopIII stars to the total metal content of each cell independently.

%Finally, as part of this \dissertation we additionally produced a significant set of companion simulations at $2 \times$ lower resolution (3.6~pc instead of 1.8~pc) turning on and off each independent feedback process (ionizing radiation, radiation pressure, LW and FUV radiation, and supernovae). With these runs, we sought to build upon recent works examining the various effects of each stellar feedback process in driving the evolution of low mass dwarf galaxies \citep{Hu2016,Hu2017,Hu2018,Forbes2016}. In future work we will examine these completed runs in more detail to better understand the complex, interconnected role multi-channel stellar feedback plays in galactic evolution. With our simulations, we can place greater emphasis on how each feedback channel drives the mixing and ejection of individual metals within and beyond low mass dwarf galaxies.

@@

In addition, this work can be readily extended to further investigate how feedback governs the evolution of dwarf galaxies. We have already produced a suite of comparison simulations at a factor two lower resolution (3.6~pc) than our fiducial resolution, turning on and off each individual feedback process to understand the significance and characterize the mechanism by which each operates to regulate star formation and drive galactic winds. The feedback tested here includes supernovae, stellar ionizing radiation, photoelectric heating, Lyman-Werner radiation, and radiation pressure. The suite contains simulations including ``all-but-one" of each process, in addition to several combinations where multiple effects are excluded, and a set where the effectiveness of radiation pressure is artificially increased and decreased. Comparing the evolution of these dwarfs to our fiducial simulations presented in this \dissertation will improve our understanding of these processes, building upon previous works that have conducted similar studies in more massive dwarf galaxies \citep{Hu2016,Hu2017,Forbes2016,Hopkins2018}. Our unique contribution to this field will be in examining how these processes change how metals are retained / ejected by galactic winds and mix within the ISM.

Finally, in ongoing work we are implementing these methods in zoom-in cosmological simulations of UFD galaxy evolution. Our detailed model tracking the feedback and yields from individual stars will build substantially upon recent works investigating metal enrichment at high redshift and in UFDs and their progenitors \citep[e.g.][]{Ritter2015,Jeon2017,Corlies2018,Wheeler2018,Agertz2019}. This work includes additional improvements to the simulations presented in this \Dissertation, including a model for Pop III star formation and chemical enrichment adapted from \cite{Wise2012a}, a new IMF sampling scheme that allows us to follow very low mass stars ($M_* \lesssim 2$~M$_{\odot}$) -- which have no significant feedback or metal enrichment, but are important long-lived tracers of galactic chemical enrichment -- aggregated together into a single particle during each star formation event, improvements to the numerical implementation to ease memory consumption in these significantly larger simulations, and new metal tracer fields to follow the contribution of AGB, core collapse SNe, Type Ia SNe, and PopIII stars to the total metal content of each cell independently. This will provide a more direct, fundamental test of the underlying physics that governs galactic evolution and produce simulated galaxies that can be more readily compared to observations of nearby dwarf galaxies.
